### 2018-04-10 latex/icpp18/introduction.tex
spmm dense introduction

There is a lack of highly scalable parallel algorithms and software for computing sparse NMF.
MPI-FAUN \cite{KBP17} software framework enables computing the NMF for dense and sparse data, yet the algorithm is optimized for dense input matrices and employs the same communication scheme to carry out sparse NMF computations.
It uses a regular, Cartesian partitioning of the input matrix $\AA$ across processors that is oblivious to the nonzero pattern.
The advantages of this approach are that the resulting factor matrix communication (to compute the SpMMs) is also regular and cast as low-latency MPI collective operations and that the rest of the computation (including NLS updates) is perfectly load-balanced.
The disadvantage is that because the partition ignores the sparsity of $\AA$, the approach will often communicate more data than necessary, as some processors will receive data that they do not need for local computation.
For sparse tensor factorization, there are scalable algorithms and software resembling NMF kernels with sparse irregular computational patterns~\cite{KU15}, yet they require an expensive partitioning step and do not necessarily yield the optimal performance for NMF.
The main advantage of this approach is that it minimizes the communication cost of the SpMM step through hypergraph partitioning, and that it employs point-to-point communication scheme which communicates elements of $\WW$ and $\HH$ only to processors in need.
The disadvantages of this approach are an upfront hypergraph partitioning cost, possible load imbalance~(particularly in the NLS 
computations) and communication imbalance. We would like to iterate that in the case of machine learning workflow, the 
computations after SpMM must deal with dense matrices, as Sparse-Dense matrix multiplication will result in a dense matrix. 
Hence, the partitioning strategy considering only SpMM will result in a huge imbalance and in this specific case of NMF on NLS computations. 
Our goal is in this paper is to fill in this gap between two approaches by comparing and evaluating them in the context of NMF, then proposing a synthesis involving parallel algorithms as well as communication and partitioning schemes enabling scalability to thousands of processes.

>>>>

There is a lack of highly scalable parallel algorithms and software for computing sparse NMF.
MPI-FAUN \cite{KBP17} software framework enables computing the NMF for dense and sparse data, yet the algorithm is optimized for dense input matrices and employs the same communication scheme to carry out sparse NMF computations.
It uses a regular, Cartesian partitioning of the input matrix $\AA$ across processors that is oblivious to the nonzero pattern.
The advantages of this approach are that the resulting factor matrix communication (to compute the SpMMs) is also regular and cast as low-latency MPI collective operations and that the rest of the computation (including NLS updates) is perfectly load-balanced.
The disadvantage is that because the partition ignores the sparsity of $\AA$, the approach will often communicate more data than necessary, as some processors will receive data that they do not need for local computation.
For sparse tensor factorization, there are scalable algorithms and software resembling NMF kernels with sparse irregular computational patterns~\cite{KU15}, yet they require an expensive partitioning step and do not necessarily yield the optimal performance for NMF.
The main advantage of this approach is that it minimizes the communication cost of the SpMM step through hypergraph partitioning, and that it employs point-to-point communication scheme which communicates elements of $\WW$ and $\HH$ only to processors in need.
The disadvantages of this approach are an upfront hypergraph partitioning cost, possible load imbalance~(particularly in the NLS 
computations) and communication imbalance. We would like to iterate that in the case of machine learning workflow, the 
computations after SpMM must deal with dense matrices, as Sparse-Dense matrix multiplication will result in a dense matrix. 
Hence, the partitioning strategy considering only SpMM will result in a huge imbalance in the overall workflow especially on computations after SpMM and in this specific case of NMF on NLS computations. 
Our goal is in this paper is to fill in this gap between two approaches by comparing and evaluating them in the context of NMF, then proposing a synthesis involving parallel algorithms as well as communication and partitioning schemes enabling scalability to thousands of processes.
